\section{Conclusion}

Decentralized document verification protocols offer clear advantages over traditional centralized systems.
Properties such as high availability, resistance to denial-of-service attacks, and immutability are inherent to the Bitcoin network and make it a suitable foundation for long-term verification mechanisms.
However, existing approaches studied in the literature reveal that Bitcoin-based solutions have historically faced limitations in scalability, privacy, and expressiveness when applied to document verification.
Protocols such as \texttt{OP\_RETURN}-based anchoring and OpenTimestamps exhibit structural constraints that restrict their applicability to large-scale or privacy-sensitive use cases.

With the activation of the Taproot upgrade in 2021, Bitcoin introduced new cryptographic primitives that significantly expand its design space.
Despite these advancements, this work identifies that Taproot has not yet been explored in the literature as the basis for a complete decentralized document verification protocol.
This observation motivates the investigation proposed in this thesis.

The protocol presented in this work leverages Taproot’s cryptographic capabilities to support document integrity verification through Merkle tree commitments and issuer authenticity verification through Schnorr digital signatures.
It further enables document revocation through transaction chaining and supports signing key rotation without invalidating previously issued documents.
By anchoring documents in batches and avoiding explicit disclosure of auxiliary data, the protocol preserves privacy by preventing information leakage about other documents within the same batch.

This work explicitly limits its scope to cryptographic verification and does not address the semantic validity of document contents, which remains the responsibility of the issuing organization.
Additionally, long-term preservation of documents and validation proofs after issuance is not considered.
Future work includes the study of archival mechanisms for long-term proof persistence, enhanced revocation strategies, and incentive or penalty schemes for entities that attempt to present invalid or revoked documents as valid.

\subsection{Contribution}

This proposal introduces an original protocol for decentralized document verification built natively on Taproot over the Bitcoin network. Unlike previous approaches, the protocol leverages Taproot and MAST not as an isolated anchoring mechanism, but as a complete cryptographic verification architecture.

By publishing a single P2TR output, the system enables autonomous verification of document integrity, issuer authenticity, and revocation status for a batch of documents, while preserving the privacy of both the issuer and the end user. Verification is performed entirely off-chain, without relying on smart contracts, external infrastructures, or third-party services.

Additionally, the protocol is efficient in its use of block space, as it aggregates multiple documents into a single transaction and maintains an on-chain cost that remains practically constant, regardless of batch size. This combination of privacy, scalability, and operational simplicity constitutes a concrete contribution to the design of document verification systems on the Bitcoin base layer.


