\section{Introduction}

Certificates issued by an organization are proof of the occurrence of a fact.
The verification that this document has not been forged, manipulated, or revoked depends on the issuer.
Confidence in the organization rests on this procedure being efficient and accessible.

The verification of a certificate has historically been done in the physical world in a centralized manner.
This entails significant costs to maintain high availability and efficiency.
With the arrival of the Internet and the proliferation of computers, it was inevitable to bring both issuance and verification into a centralized digital space.

Bitcoin began operating in 2009 as the first permissionless decentralized network.
It has not had any censorship events or downtime as of the writing of this document.
It allows the execution of limited smart contracts following the UTXO model, as well as data anchoring.
Limited by design to reduce the attack surface for denial-of-service and due to the high cost of anchoring bytes of information, it was not considered a scalable solution for document verification.

Various studies and proposals have turned to using other blockchain networks, such as Ethereum.
By allowing the execution of Turing-complete smart contracts, they facilitate the validation and verification of certificates.
However, they introduce different trust assumptions and greater complexity in consensus, which increases the risk of censorship or rollbacks, compromising verification autonomy.

Bitcoin has a protocol that has evolved over its 16 years of existence.
It had two major upgrades accepted by the community, Segwit (2017) and Taproot (2021).
These improvements in security, privacy, and blockspace efficiency allow for the expansion of use cases.
Utilizing the cryptographic tools enabled by Taproot, we propose a methodology for document verification directly on the base layer of Bitcoin.
