\section{Literature Review}

A Bitcoin transaction is a variable-sized string of Bytes with a theoretical maximum of 1 MB.
It has information sections such as the timestamp, version, outputs, etc.
We highlight one section in particular: the \texttt{scriptPubKey} in the output.
The type of transaction is configured depending on the type of \texttt{scriptPubKey}.

The standard ones being Pay-to-PublicKeyHash (P2PKH) and Pay-to-ScriptHash (P2SH), known as legacy transactions.
These transactions suffer from a vulnerability called malleability, which was resolved with the arrival of SegWit in 2017.
Malleability is a vulnerability that allows the DER encoding of an ECDSA digital signature to be modified without invalidating it, thereby changing the transaction index (txid).

This vulnerability was the primary motivation for developing SegWit in 2017.
It modifies the transaction structure by moving the signature information (the \texttt{scriptSig}) to a separate section called the Witness.
This introduced new transaction types: Pay-to-Witness-PublicKeyHash (P2WPKH) and Pay-to-Witness-Script-Hash (P2WSH).

\subsubsection{Data Anchoring Techniques}

The use of Bitcoin was initially limited only to the peer-to-peer transmission of monetary value.
By maintaining a distributed and consistent ledger across the entire network, the possibility of using it for other applications soon became clear.
In the work 'Trusting records: is Blockchain technology the answer?'~\cite{Lemieux2016Trusting}, it is studied whether this technology is appropriate for preserving the integrity of digital information using various ISO standards.
The analysis concludes that, while Blockchain is useful for addressing information integrity in the short term, it does not guarantee the initial reliability of the records and presents several limitations as a long-term solution for authentic digital preservation.
The main argument is the vulnerability of digital signature algorithms and hashes like sha256 and ripemd160.

With the exception of Coinbase transactions (where new Bitcoin tokens are issued in each block), Bitcoin transactions do not contain a default space for inserting arbitrary information.
The paper 'Decentralized Trusted Timestamping Blockchain'~\cite{Gao2017Decentralized} proposes a method to introduce metadata within P2PKH transactions.
It encodes information inside a "fake" address, losing control by not having the corresponding private key (sk).
By being able to introduce $N$ outputs in a transaction, the proposed method has $20 \times N$ available bytes.
This method permanently burns the Bitcoin in each output, making it one of the most expensive.

In the work 'Data Insertion in Bitcoin’s Blockchain'~\cite{Sward2018DataInsertion}, an exhaustive search is made of the ways used to insert data into transactions.
It identifies two categories: transactions that use ScriptHash and those that use "fake addresses" as in the case described in the previous paper.
It finds that the use of ScriptHashes is a more complex process to perform.
It has advantages in not being limited to 20 bytes; however, its construction is non-standard and needs to be developed at each stage.
It concludes that "no universally optimal method exists" as there is a trade-off between the amount of data and the anchoring costs.
It also mentions the need for studies to investigate the impact of SegWit, as it is limited to analyzing legacy transactions.

The paper "A Journey into Bitcoin Metadata"~\cite{Bartoletti2019BitcoinMetadata} also analyzes techniques for anchoring metadata in Bitcoin and its applications, such as proof of ownership and financial records.
It explores in detail whether the different anchoring methods negatively affect Bitcoin's primary function.
Finding that one problem is the bloat of the Unspent Transaction Output (UTXO) set.
In this regard, it highlights that methods using \texttt{OP\_RETURN} are those that generate the least impact, occupying less space in relation to the total size of the blockchain.

\texttt{OP\_RETURN} methods utilize a transaction output, without spending Bitcoin.
It is a \texttt{scriptPubKey} that contains the \texttt{OP\_RETURN} opcode, which terminates the script execution.
The work "Dominating OP Returns: The Impact of Omni and Veriblock on Bitcoin"~\cite{Strehle2020DominatingOPReturns} analyzes 32.4 million transactions between 2018 and 2019, finding that the use of \texttt{OP\_RETURN} comes mainly from 37 services.
Primarily: Omni, which anchors financial information related to the use of USDT, and Veriblock, which uses it for the security of other blockchains using Proof-of-Proof.
Showing that \texttt{OP\_RETURN} was established as the standard and dominant method for embedding non-monetary data in the Bitcoin blockchain.
It measures their impact on the Bitcoin network, finding that the analyzed transactions contribute approximately 10 GB to the network despite being limited to 80 bytes per output.

The use of legacy transactions for data anchoring has two problems in its design: lack of privacy and the malleability vulnerability.
When analyzing the blockchain, the \texttt{OP\_RETURN} content is clearly identified, losing the privacy of the anchored information.
Malleability is a vulnerability that allows an attacker to modify the encoding of a digital signature without affecting its validity, creating two valid transactions but with different transaction IDs (txid).

These two problems are the main reason for the implementation of BIP141 SegWit.
In the paper 'Analysis of segregated witness implementation...'~\cite{Kedziora2022AnalysisSegWit}, the authors study the impact after the softfork's application.
It finds benefits in separating the digital signature from the transaction's data block into another space called the Witness: efficiency and security.
A SegWit transaction is 75\% less costly and solves the malleability problem.
It finds that data anchoring via the Witness expands the limitation in Bytes.
While they do not affect the bloating of the UTXO set, they need to be fully declared and spent to be anchored to the Blockchain.
Concluding that the implementation was successful and opens the door to new types of applications with more versatile Smart Contracts.

The development of new cryptographic tools is fundamental to improving performance in Blockchain networks.
Studies like Gayoso~\cite{Gayoso2020AnalysisCryptographic}, Fang~\cite{Fang2020DigitalSignature}, and Zhang~\cite{Zhang2025HighEfficiencySchnorrMultiSignature} analyze the essential technological pillars: hash functions, digital signatures, and Merkle trees.
They suggest the use of new digital signature schemes, highlighting Schnorr signatures.
Adding privacy and efficiency in validation algorithms, they show their advantage in aggregate, group, and threshold signature schemes.
The authors agree that Schnorr is a scheme that meets the non-repudiation requirements in Bitcoin.

These foundations are the basis for a new softfork proposed in BIP341 Taproot; it is an extension of SegWit, considered a new version of it.
It proposes a new transaction type, P2TR (Pay-to-TapRoot).
Since its activation in November 2021, Taproot has been successfully implemented.
Opening the possibility for new applications.

\subsubsection{Alternative Systems and Applications (Ethereum)}

It has been shown that data can be anchored in Bitcoin in different ways.
The limitation in Bitcoin's Script execution also limits the applications built on top of it.
To overcome these limitations, new Blockchain networks have been created that change the UTXO design that Bitcoin uses.
The one with the most development is Ethereum, which emulates a virtual machine called the EVM.

Different works such as Wang~\cite{Wang2019BlockchainCT}, Li~\cite{Li2020DecentralizedPKI}, Pennino~\cite{Pennino2021EfficientCertification}, Palmisano~\cite{Palmisano2022NotarizationAntiPlagiarism}, Mishra~\cite{Mishra2025BlockchainDecentralized}, and Harer~\cite{Harer2022DecentralizedAttestation} make use of Smart Contracts to attest information such as certificates, dates, and XML structures, respectively.
Sophisticated certification and attestation in Smart Contract-based architectures depend on costly execution (gas fees) and programmable complexity (Solidity, EVM) to manage the logic of certificate verification, identity, and permissions.
The applications vary, but they all make use of data anchoring.

A case of special interest are the works related to the institutional use of document verification.
In 'Blockchain and Higher Education Diplomas'~\cite{Castro2021BlockchainHigherEducation}, the need to eradicate fraud and falsification of diplomas and the sale of fraudulent certificates, a global market estimated at 2 billion USD, is developed.
The study notes that, as a radical innovation, blockchain adoption must overcome technical barriers (scalability, error correction, complexity) and cultural/social barriers (lack of standards, resistance to change).

'A Systematic Literature Review on Blockchain-Based Systems for Academic Certificate Verification'~\cite{Rustemi2023SLRBlockchainCertificates} identifies that most existing platforms are implemented on the Ethereum blockchain and services are executed via Smart Contracts.
The immutability of Smart Contracts is a challenge, as the revocation of a diploma implies a change in the hash value, which challenges the student's identity and is an inherently difficult or impossible process in the blockchain system.
Verification is the only process that can be executed completely without third-party intervention, while generation and revocation remain challenging.

In summary, the literature presents a clear trade-off: legacy Bitcoin solutions (like \texttt{OP\_RETURN}) lack privacy and scalability, while Smart Contract solutions (like Ethereum) are computationally expensive.

Taproot, with its efficient Schnorr signatures and MAST (Merkleized Abstract Syntax Trees), allows encoding complex verification logic (similar to an SC), but does so with a minimal transaction footprint on the Bitcoin blockchain, avoiding the high operational cost of Ethereum's EVM and overcoming the data capacity limitations of old Bitcoin methods like \texttt{OP\_RETURN}.