\section{Motivation}

The primary motivation for this work arises from the fragility of centralized document verification models.
When an issuing organization becomes unavailable, ceases operations, or loses its internal records, the ability to verify previously issued documents is effectively lost.
This dependency introduces a single point of failure that undermines long-term verifiability and trust.

In contrast, the reliability of a decentralized system is rooted in its consensus protocol.
Bitcoin operates under a Proof-of-Work (PoW) consensus mechanism, providing strong security guarantees, immutability, and resistance to denial-of-service (DoS) attacks.
These properties make Bitcoin particularly suitable for applications that require long-term availability and independent verification.

Other blockchain-based solutions for document verification often rely on alternative networks such as Ethereum.
While these platforms benefit from Turing-complete smart contracts and flexible execution models, they also introduce different scalability, cost, and trust trade-offs compared to Bitcoin.
In many cases, verification logic is executed on-chain, increasing operational complexity and dependency on the continued availability of smart contract infrastructure.

The activation of Taproot significantly improves Bitcoin’s privacy and block space efficiency at the base layer.
These advancements enable new application designs that were previously impractical on Bitcoin.
Motivated by these developments, this work explores a document verification protocol that operates entirely on Bitcoin’s main layer, without requiring additional protocol upgrades, second-layer solutions, or reliance on external blockchain networks.
