\section{Problem Statement}

Existing decentralized approaches to document verification exhibit limitations that depend strongly on the underlying blockchain infrastructure.
In the context of Bitcoin, current solutions based on protocols such as OpenTimestamps, Blockcerts, and \texttt{OP\_RETURN} either expose embedded metadata, compromising privacy, or incur high costs in terms of block space and transaction fees.
As a result, these approaches do not scale efficiently for large batches of documents or long-term institutional use.

The Taproot upgrade introduces new cryptographic primitives that significantly improve efficiency and privacy at the Bitcoin base layer.
By adopting Schnorr signatures, Taproot enables more compact and flexible multisignature constructions.
Additionally, the use of Merkleized Abstract Syntax Trees (MAST) allows multiple script conditions to be committed within a single transaction output, reducing on-chain footprint while preserving verifiability.
Despite these capabilities, Taproot has not yet been leveraged to design a complete protocol for decentralized document verification that simultaneously supports integrity, issuer authenticity, and revocation.

\subsection{Research Questions}

This work aims to address the following research questions:
\begin{itemize}
    \item How can Taproot be used to build a decentralized document verification protocol on Bitcoin?
    \item How does using Merkle tree commitments in Taproot improve privacy and scalability during document verification?
    \item How can document revocation be supported within a Taproot-based verification protocol?
\end{itemize}

\subsection{Specific Objectives}

To answer the proposed research questions, the following objectives are defined:
\begin{enumerate}[label=\alph*)]
    \item To design a Merkle tree structure that enables the verification of document integrity.
    \item To design a mechanism for verifying issuer authenticity using digital signatures.
    \item To design a revocation tracking mechanism that allows the verification of document validity.
    \item To propose a proof of concept (PoC) of the proposed protocol on a Bitcoin test network.
    \item To estimate the protocol performance in terms of scalability, computational complexity, and on-chain costs measured in virtual bytes (vBytes).
\end{enumerate}
