\begin{abstract}

Document verification is a fundamental mechanism for detecting document falsification and fraud in institutional processes.
Traditional approaches are based on centralized infrastructures, which incur high operational costs, limited availability, and single points of failure.
When the issuing organization is compromised or becomes unavailable, verification capability is effectively lost.

Bitcoin, introduced in 2009, operates as a decentralized network secured by a proof-of-work consensus protocol.
Its properties of immutability, high availability, and permissionless access make it a suitable foundation for applications beyond monetary use.

Bitcoin has been widely used for anchoring non-repudiable information.
Several data anchoring techniques have been proposed for Bitcoin, including OP\_RETURN, fake addresses, OpenTimestamps, and Script- and Witness-based methods.
However, these approaches exhibit limitations in scalability, privacy, and usability.
The activation of the Taproot upgrade in 2021 introduced more efficient and private cryptographic primitives, enabling new application designs at the base layer.

This work proposes a novel protocol for decentralized document verification using Taproot.
The protocol supports integrity verification, issuer authenticity verification, and revocation status checking.
By anchoring a batch of $N$ documents to a single pay-to-taproot P2TR output, the system improves scalability and preserves privacy.
Verification is performed entirely off-chain, and the protocol evaluates both computational costs and on-chain block space usage.
\end{abstract}